Blaming the Typeless: \\ Scalable, Human-Centric Python Localization

### 2018-08-22 paper/main.tex
more diversity info

In Nate there were 20 different functions. All the programs were doing the
same 20 functions.

Here, people are doing anything they want. Random stuff going on over here.

We want to show that we get high accuracy *over a heterogenous set of
programs*. Leon?

prereq: cluster the programs based on similarity. Using agglomerative
clustering plus python edit distance. How good the performance will be?
right now it creates a clustering tree. binary tree, based on min distance.
can give it a threshhold for distance (elements w/in a group). don't forget
to rename all variables to X, perhaps normalize (edit distance size).

leon: could also cluster by inconsistency coefficient

PLOT Type: line graph (standard plot)

PLOT X AXIS: clustering distance parameter

PLOT Y AXIS: number of different clusters found

PLOT LINES: line 1 is ``Nate programs'', line 2 is ``25\%, at random, of our dataset''

Goal: show that ``no matter how you decide program differences'', our
dataset is always more diverse/heterogeneous than is previous work. So our
success is amazing.

@@

\begin{figure}
\begin{tikzpicture}
  \begin{axis}[
      xlabel=Clustering threshold parameter,
      ylabel=\# of naturally occuring program clusters,
      xmax=1.25,
      legend style={at={(0.5,0.25)},anchor=north}
    %   ytick distance=1000,
    %   title=Linkage: single
    ]
    \foreach \language in {ocaml, python}
    {
      \addplot+[mark=none] table[x=threshold, y=\language_cluster_counts_single, col sep=comma]{cluster_counts.csv};
    }
\legend{NATE OCaml data, Our Python data}
\end{axis}
\end{tikzpicture}
\caption{Number of clusters when data is clustered by single linkage.
  A linkage tree is built using agglomerative clustering, and the resulting tree is broken up
  into clusters by thresholding the inconsistency coefficient. The inconsistency coefficient
  of a link is its distance, z-scored against other nearby links. TODO: find a way to include
  the Rosetta Code baseline in this graph? better not to devote multiple graphs to this issue.}
\label{fig:diversity}
\end{figure}

@@

\bibliographystyle{abbrv}
\bibliography{nanomaly,sw,temp,slice}

>>>>

%In Nate there were 20 different functions. All the programs were doing the
%same 20 functions.
%
%Here, people are doing anything they want. Random stuff going on over here.
%
%We want to show that we get high accuracy *over a heterogenous set of
%programs*. Leon?
%
%prereq: cluster the programs based on similarity. Using agglomerative
%clustering plus python edit distance. How good the performance will be?
%right now it creates a clustering tree. binary tree, based on min distance.
%can give it a threshhold for distance (elements w/in a group). don't forget
%to rename all variables to X, perhaps normalize (edit distance size).
%
%leon: could also cluster by inconsistency coefficient
%
%PLOT Type: line graph (standard plot)
%
%PLOT X AXIS: clustering distance parameter
%
%PLOT Y AXIS: number of different clusters found
%
%PLOT LINES: line 1 is ``Nate programs'', line 2 is ``25\%, at random, of our dataset''
%
%Goal: show that ``no matter how you decide program differences'', our
%dataset is always more diverse/heterogeneous than is previous work. So our
%success is amazing.

@@

\begin{figure}
\begin{tikzpicture}
  \begin{axis}[
      xlabel=Clustering threshold parameter,
      ylabel=\# of naturally occuring program clusters,
      xmax=1.25,
      legend style={at={(0.5,0.25)},anchor=north}
    %   ytick distance=1000,
    %   title=Linkage: single
    ]
    \foreach \language in {ocaml, python}
    {
      \addplot+[mark=none] table[x=threshold, y=\language_cluster_counts_single, col sep=comma]{cluster_counts.csv};
    }
\legend{NATE OCaml data, Our Python data}
\end{axis}
\end{tikzpicture}
\caption{Number of clusters when data is clustered by single linkage.
  A linkage tree is built using agglomerative clustering, and the resulting tree is broken up
  into clusters by thresholding the inconsistency coefficient. The inconsistency coefficient
  of a link is its distance, z-scored against other nearby links. TODO: find a way to include
  the Rosetta Code baseline in this graph? better not to devote multiple graphs to this issue.
  LM: the y axis could be made into a percentage of the dataset size, but that might be
  weird?}
\label{fig-diversity}
\end{figure}

As a measure of the diversity of our dataset, we clustered the programs in
our dataset and the NATE dataset, then compared the number of resultant clusters.
To cluster programs, we transform the program ASTs into flat strings, and then
perform agglomerative clustering on them, with string edit distance as a distance
metric.

\subsubsection{Transformation and Distance Metric}

To measure the distance between programs, we flattened their ASTs into strings
of tokens, and then computed the Levenshtein edit distance~\cite{levenshtein}.
The decision to flatten the tree rather than compute a distance directly on the
the ASTs is practical; the state of the art in tree edit distance has cubic
asymptotic complexity~\cite{PAWLIK2016157}, while the levenshtein distance can be
computed in time quadratic in the length of the strings~\cite{lev-quadratic}.
When we attempted a test run using tree edit distance as the metric, it did not
complete and we determined that it would not be practical to use.

The flattening process, which is based on \cite{dist-site}, is defined
recursively: the string for an AST
begins with a unique token corresponding to the AST node type, which is
followed by the concatenation of the transformations of the tree's subtrees.
The string ends with an ``end'' token to denote the end of the tree. This
preserves the tree structure of the AST.

\subsubsection{Clustering Algorithm}

We performed agglomerative clustering on the programs~\cite{modern-clustering}.
This is a method of clustering in which every datapoint starts in its own
cluster, and the two closest clusters are merged every step of the algorithm,
until every datapoint is in the same cluster. This produces a binary tree, where
each internal node corresponds to a cluster formed by merging two other clusters,
and each leaf corresponds to an item in the dataset. We used single linkage,
which means that the distance between clusters of multiple items was considered
to be the minimum distance between their elements.

This alone produces a tree rather than a list of disjoint clusters, which makes
different clusterings difficult to compare. To produce a cluster count, we used
the SciPy function \lt{scipy.cluster.hierarchy.fcluster}, which forms disjoint
clusters by splitting the tree at internal nodes where the two joined clusters
are more distant than some threshold. We used the \lt{'inconsistent'} method with
depth set to 2, which specifies that the distances at each node should be
z-scored against the distances at other nodes at the same level in the tree, and
any nodes 1 or 2 levels below~\cite{inconsistent}. This adapts to differences in
scale of edit distance that might arise between python and OCaml due to
programming language expressivity or other factors, making cluster counts at each
threshold value comparable.

\figref{diversity} shows the results of clustering on our dataset and the NATE
dataset. For all values of the inconsistency threshold, the algorithm computed a
higher cluster count for the Python data, suggesting that it contains a more
diverse set of programs than the OCaml dataset. The similarity in the graphs
appears to confirm the comparibility of cluster counts between the two datasets.

@@

\bibliographystyle{abbrv}
\bibliography{nanomaly,sw,temp,slice,cluster}
